---
title: "Competetion"
author: "Alex Federspiel"
date: "3/26/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r packages, include = FALSE}
  library(readr)
  library(tidyverse)
  library(caret)
  library(corrplot)
```

```{r data, include = FALSE}
  competition_train <- read_csv("C:/Users/afede/Desktop/Grad School/Data Science/CourseContentRepo/Assignments/competition/competition-train.csv")
```

```{r first 10 variables, include = FALSE}
  mini_competetion <- competition_train %>% select(X1:X10, outcome)
  df_mini_competetion <- data.frame(mini_competetion %>% select(- outcome))
  mini_preprocess_fit <- preProcess(df_mini_competetion, 
                                    method = c("BoxCox", "center", "scale"))
  mini_preprocess <- predict(mini_preprocess_fit, df_mini_competetion)
```

## Missing Values
No missing values.

```{r missing values, echo = FALSE, warning = FALSE}
  total_nan <- function(x) sum(is.na(x))
  mini_competetion %>% summarise(across(everything(), total_nan))
```


## X1
```{r X1, echo = FALSE}
  summary(mini_competetion$X1)
  ggplot(mini_competetion, aes(X1)) + geom_histogram(color = "white", bins = 40)
  ggplot(mini_competetion, aes(y = X1)) + geom_boxplot()
```

Majority of the values are $0$. Has near-zero variance. Should drop.

* The fraction of unique values over the sample size is: $\frac{315}{5000}=6.3\%$. Textbook says it should be low like $10\%$.

* Ratio of frequency of most prevalent value to frequency of second most is: $\frac{2969}{91}=32.62$. Textbook says it should be high like $20$.

```{r sum zero, echo = FALSE}
  head(mini_competetion %>% group_by(X1) %>% summarise(count = n()) %>% 
    arrange(desc(count)))
```

## X2
```{r X2}
  summary(mini_competetion$X2)
  ggplot(mini_competetion, aes(X2)) + geom_histogram(color = "white", bins = 40)
  ggplot(mini_preprocess, aes(X2)) + geom_histogram(color = "white", bins = 40)
```

Data looks good after preprocess.

## X3
```{r x3}
  summary(mini_competetion$X3)
  ggplot(mini_competetion, aes(X3)) + geom_histogram(color = "white", bins = 40)
  ggplot(mini_preprocess, aes(X3)) + geom_histogram(color = "white", bins = 40)
```

Data looks good after preprocess.

## X4
```{r x4}
  summary(mini_competetion$X4)
  ggplot(mini_competetion, aes(X4)) + geom_histogram(color = "white", bins = 40)
  ggplot(mini_preprocess, aes(X4)) + geom_histogram(color = "white", bins = 40)
```

Data looks good after preprocess.

## X5
```{r x5}
  summary(mini_competetion$X5)
  ggplot(mini_competetion, aes(X5)) + geom_histogram(color = "white", bins = 40)
  ggplot(mini_preprocess, aes(X5)) + geom_histogram(color = "white", bins = 40)
```

Data looks good after preprocess.

## X6
```{r x6}
  summary(mini_competetion$X6)
  ggplot(mini_competetion, aes(X6)) + geom_histogram(color = "white", bins = 40)
  ggplot(mini_competetion, aes(y = X6)) + geom_boxplot()
  ggplot(mini_preprocess, aes(X6)) + geom_histogram(color = "white", bins = 40) +
    labs(title = "Preprocess")
```

Not sure if we should remove or not. Looks skewed even after preprocess. The variance looks low as well.

* Unique values: $\frac{204}{5000}=4.08\%$
* Ratio of most common to second most: $\frac{1202}{584}=2.06$

The ratio isn't outrageous but uniqueness is small. 

## X7
```{r x7}
  summary(mini_competetion$X7)
  ggplot(mini_competetion, aes(X7)) + geom_histogram(color = "white", bins = 40)
  ggplot(mini_competetion, aes(y = X7)) + geom_boxplot()
```

Near-zero variance. Most likely should drop.

* Unique values: $\frac{320}{5000}=6.4\%$
* Common Ratio: $\frac{1224}{372}=3.29$

Again, ratio of most common to second isn't very large but the percent of unique values is low. 

## X8
```{r x8}
  summary(mini_competetion$X8)
  ggplot(mini_competetion, aes(X8)) + geom_histogram(color = "white", bins = 40)
  ggplot(mini_competetion, aes(y = X8)) + geom_boxplot()
```

Another near-zero variance. Most values are $0$.

* Uniqueness: $\frac{172}{5000}=3.44\%$
* Common Ratio: $\frac{1642}{934}=1.76$

Probably drop.
```{r, include = FALSE}
  mini_competetion %>% group_by(X8) %>% summarise(count = n()) %>% 
    arrange(desc(count))
```


## X9
```{r x9}
  summary(mini_competetion$X9)
  ggplot(mini_competetion, aes(X9)) + geom_histogram(color = "white", bins = 40)
  ggplot(mini_preprocess, aes(X9)) + geom_histogram(color = "white", bins = 40)
```

Data looks good after preprocess.

## X10
```{r x10}
  summary(mini_competetion$X10)
  ggplot(mini_competetion, aes(X10)) + geom_histogram(color = "white", bins = 40)
  ggplot(mini_preprocess, aes(X10)) + geom_histogram(color = "white", bins = 40)
```

```{r, include = FALSE}
    mini_competetion %>% group_by(X10) %>% summarise(count = n()) %>% 
    arrange(desc(count))
```

## Summary
* Remove: X1
* Potential Removals: X6, X7, X10
  + These have skew and outliers. Not entirely sure what we can do about these.
* Removed Predictors from Cutoff Matrix: X11, X12, X16, X17, X18

```{r}
  correlation_matrix <- cor(competition_train %>% select(where(is.numeric)))
  corrplot(correlation_matrix, order = "hclust")
  
  predictors_only <- data.frame(competition_train %>% select(-outcome))
  predictors_only_correlation_matrix <- cor(predictors_only)
  cutoff <- 0.90
  
  names_of_predictors_to_remove <- findCorrelation(predictors_only_correlation_matrix, 
                                                 names = TRUE, cutoff = cutoff)
  removed_predictors <- predictors_only %>% 
    select(-all_of(names_of_predictors_to_remove))
```
